Using Energy Smart Meter Data and Socio-Demographic Indicators to Understand Consumption in Relation to Deprivation Levels

### 2018-06-02 paper/Paper_1_v02.Rmd
more edits
further edits to draft

On the international scale, the increased release of greenhouse gases into the atmosphere has caused an rise in global temperatures of 0.8&deg;C  since 1880 - but two thirds of this warming has occured since 1975 [@Observatory]. According to the IEA (International Energy Association), energy efficiency can contribute up to 49% of the energy related CO2 emission reductions that are nedded to limit global temperature increases to less than 2&deg;C by 2050. "Energy efficiency, including energy conversation, is a long-term priority for G20 members, as it constitures the optimum utilisation of energy resources" [G20 https://ec.europa.eu/energy/sites/ener/files/documents/G20%20Energy%20Efficiency%20Leading%20Programme.pdf] (pp.4). G20 members account for over 80% of both global energy consumption and greenhouse gases but have proven experience in "energy efficiency measures and in achieving energy reductions". From 1990 to 2013,the G20s total energy consumption savings reached about 4.3 billion TOE (define TOE) and about 10.4 billion tonnes of carbon dioxide emissions were avoided. Energy efficiency is one of the most important mechanisms through which countries can act to mitigate climate change in the short to long term.

@@

# The dichotomy between energy and fuel poverty

@@

According to the UK government, a house is considered to be in fuel poverty if "They have required fuel costs that are above average **and** if they were to spend that amount they would be left with a residual income below the official poverty line." It is stated that the main drivers are "household incomes, household energy efficiency and fuel prices". This definition of fuel poverty has important policy implications; for the formulation of, determining it's scale and nature, targeting a strategy and monitoring progress [@Moore2012]. @Boardman1991 was the first to give a formal definiton in her book *Fuel Poverty*. This is the first mention of the 10% threshold, which has become commonplace for policymakers when defining the fuel poor, and is the basis for the current definition used by the DECC and other major stakeholders [@Liddell2012].

@@

To get a deeper understanding of the demographic characteristics of Smart Meter users, census data was appended to the energy usage data. This had to be reweighted from output area level to be reapportioned to postcode sector level. This was done using the postcode headcount dataset, and an output area to postcode sector lookup table for calculating proportions. (This method is preferable to the GIS method of creating a shapefile based on overlapping boundaries as this is prone to inaccuracies such as slither polygons and is computationally much more intensive).

@@

The literature time and again refers to accomodation type as a major factor in deprivation levels. The deprivation model ran here included the proportion of accomodation types in each PCS; detached, semi-detached, terrace, flat or shared accomodation. Due to the size of two of the groups, flats and shared accommodation were combined. As the largest category overall, semi-detached was used as the reference category.

A daily temporal profile was calculated showing total energy usage per day; these were combined with the PCS townsend scores and accomodation types and a linear regression model ran.

The deprivation model explained R^2^ = 15.54% (adjusted R2 = 15.49%) of the variability in domestic energy consumption, p < 2.2e-16.

A detached and typically larger house than the reference category (semi-detached) would use on average an extra 1953.2 kW per day in comparison, whereas a smaller terrace house would use less to the sum of 135.5 kW per day. Detached houses typically have a greater number of rooms and lower levels of overcrowding, therefor more rooms need heating and lighting (which account for X amount of average consumption). This model also suggests that as Townsend scores increase, energy usage decreases, which as a measure of deprivation is to be expected – the higher the level of deprivation the less likely people are to be able to afford to be wasteful and consume unnecessary energy. They are also likely to live in smaller houses.

>>>>

On the international scale, the increased release of greenhouse gases into the atmosphere has caused an rise in global temperatures of 0.8&deg;C  since 1880 - but two thirds of this warming has occured since 1975 [@Observatory]. According to the IEA (International Energy Association), energy efficiency can contribute up to 49% of the energy related CO2 emission reductions that are nedded to limit global temperature increases to less than 2&deg;C by 2050. "Energy efficiency, including energy conversation, is a long-term priority for G20 members, as it constitures the optimum utilisation of energy resources" [@G20-2020] https://ec.europa.eu/energy/sites/ener/files/documents/G20%20Energy%20Efficiency%20Leading%20Programme.pdf] (pp.4). G20 members account for over 80% of both global energy consumption and greenhouse gases but have proven experience in "energy efficiency measures and in achieving energy reductions". From 1990 to 2013,the G20s total energy consumption savings reached about 4.3 billion TOE (define TOE) and about 10.4 billion tonnes of carbon dioxide emissions were avoided. Energy efficiency is one of the most important mechanisms through which countries can act to mitigate climate change in the short to long term.

@@

# Policy impacts from the dichotomy between energy poverty and fuel poverty

@@

According to the UK government, a house is considered to be in fuel poverty if "They have required fuel costs that are above average **and** if they were to spend that amount they would be left with a residual income below the official poverty line." It is stated that the main drivers are "household incomes, household energy efficiency and fuel prices" [@DECC2017]. This definition of fuel poverty has important policy implications; for the formulation of, determining it's scale and nature, targeting a strategy and monitoring progress [@Moore2012]. @Boardman1991 was the first to give a formal definiton in her book *Fuel Poverty*. This is the first mention of the 10% threshold, which has become commonplace for policymakers when defining the fuel poor, and is the basis for the current definition used by the DECC and other major stakeholders [@Liddell2012].

Fuel poverty in England is measured using the Low Income High Cost indicator, which as a dual indicator allows the measurement of both the extent of the problem and the depth of the problem [@DECC2017].

@@

To get a deeper understanding of the demographic characteristics of Smart Meter users, census data was appended to the energy usage data. It had to be reweighted from output area level to be reapportioned to postcode sector level which was done achieved the postcode headcount dataset, and an output area to postcode sector lookup table for calculating proportions. (Testing revealed this method is preferable to the GIS method of creating a shapefile based on overlapping boundaries as this is prone to inaccuracies such as slither polygons and is computationally much more intensive).

@@

The literature time and again refers to accomodation type as a major factor in energy consumption. This is because the type of accommodation a person lives is a good proxy for their usage habits; for example those who live alone will consume much less energy than a family with three children for a multitude of reasons; a large family are more likely to occupy a home with a larger floor space as their need is greater. They therefore have more rooms which require heating and lighting, and they will likely have a higher number of appliances, and or use these appliances such as the dishwsher or washing machine much more frequently than a single person living alone to achieve the same level of comfort.

It is also true to say that the UKs housing stock is currently much older than that in the rest of Europe. This goes some way to explaining the relative higher consumption but also at an individual level, post-war terrace houses are still extremely common, but are poorly insulated and in need of updating. Those families living in large detached and semi-detached houses are also more likely to be owner occupiers, and the DECC stated in their latest report that it is the disconnect between tenants and landlords that stands in the way of privately rented homes becoming more energy efficient, because it is unclear who should pay for and benefit from the efficiency measures. Private renters are the most likely to be in the deepest fuel poverty.

Accommodation type is also indicative

The model ran attempts to explain the energy consumption by levels of deprivation.

The deprivation model ran here included the proportion of accomodation types in each PCS; detached, semi-detached, terrace, flat or shared accomodation. Due to the size of two of the groups, flats and shared accommodation were combined. As the largest category overall, semi-detached was used as the reference category. As can be seen in the OAC group pen profiles, deprivation and housing type are closely linke so it was important to control for it to avoid the Townsend score returning a significant result as it is being used as a proxy for type of accomodation.

A daily temporal profile was calculated showing total energy usage per day; these were combined with the PCS townsend scores and accomodation types and a linear regression model ran.The energy consumption model explained R^2^ = 15.54% (adjusted R2 = 15.49%) of the variability in domestic energy consumption, p < 2.2e-16, suggesting that these results have not occured randomly.

A detached and typically larger house than the reference category (semi-detached) would use on average an extra 1953.2 kW per day in comparison, whereas a smaller terrace house would use less to the sum of 135.5 kW per day. Detached houses typically have a greater number of rooms and lower levels of overcrowding, therefor more rooms need heating and lighting (which account for 60% amount of average consumption). This model also suggests that as Townsend scores increase, energy usage decreases overall, which as a measure of deprivation is to be expected – the higher the level of deprivation the less likely people are to be able to afford to be wasteful and consume unnecessary energy. They are also likely to live in smaller houses with lower heating and lighting requirement, own and use fewer appliances and have smaller families.
